---
title: "Classification-trees-or-rules"
author: "Joseph Park"
date: "4/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## data and packages

```{r}

library(class)

library(tidyverse)

library(gmodels)

library(MLmetrics)

library(C50)

#library(Rweka)
```

```{r}
particle <- read_csv("C:/Users/parkj/OneDrive/Documents/joe/ML with R/pid-5M.csv")
```
## sample (randomly) due to too much data for my laptop to process
```{r}
set.seed(47)
particles <- particle[sample(nrow(particle), 50000), ]

particles$id <- as.character(particles$id)


print(particles[complete.cases(particles) == FALSE, ])
str(as.factor(particles$id))
print(which(is.na(particles)))
print(summary(particles))
for (i in 2:7){
  print(sd(as.matrix(particles[,i])))
}


```
```{r}
for (i in 2:7){
  hist(as.matrix(particles[,i]))
}


```
## Split data

```{r}
particles_train <- particles[1:45000, 2:7]
particles_test <- particles[45001:50000, 2:7]

particles_train_labels <- particles[1:45000, 1]
particles_test_labels <- particles[45001:50000, 1]
```
## Train classifier on training data and classify test data

```{r}
#generating classification for trials to find best accuracy
accuracy = 1:30


for (i in 1:30){

  m <- C5.0(particles_train, as.factor(particles_train_labels[,1, drop = TRUE]), trials = i)


  particles_test_pred <- predict(m, particles_test)
    
  #metric
  accuracy[i] <- Accuracy(particles_test_pred, particles_test_labels[,1, drop = TRUE])

#get best accurracy and corresponding index (with respect to Laplace estimator)
  if (i > 1){
    if (accuracy[i] > bestacc){
      bestacc <- accuracy[i]
      bestk <- i
    }
  } else {
    bestacc <- accuracy[i]
    bestk <- i
  }
}
```
## Metrics

```{r}
plot(accuracy[1:30])
print(paste("The best accuracy is", bestacc))
```
```{r}

m <- C5.0(particles_train, as.factor(particles_train_labels[,1, drop = TRUE]), trials = bestk)


  particles_test_pred <- predict(m, particles_test)

CrossTable(x = particles_test_labels[,1, drop = TRUE], y = particles_test_pred, prop.chisq=FALSE)



```

```{r}
plot(m)
```
```{r}
summary(m)
```

# now with rules

## Train classifier on training data and classify test data

```{r}
#generating classification for trials to find best accuracy
accuracy = 1:30


for (i in 1:30){

  m <- C5.0(particles_train, as.factor(particles_train_labels[,1, drop = TRUE]), trials = i, rules = TRUE)


  particles_test_pred <- predict(m, particles_test)
    
  #metric
  accuracy[i] <- Accuracy(particles_test_pred, particles_test_labels[,1, drop = TRUE])

#get best accurracy and corresponding index (with respect to Laplace estimator)
  if (i > 1){
    if (accuracy[i] > bestacc){
      bestacc <- accuracy[i]
      bestk <- i
    }
  } else {
    bestacc <- accuracy[i]
    bestk <- i
  }
}
```
## Metrics

```{r}
plot(accuracy[1:30])
print(paste("The best accuracy is", bestacc))
```
```{r}

m <- C5.0(particles_train, as.factor(particles_train_labels[,1, drop = TRUE]), trials = bestk)


  particles_test_pred <- predict(m, particles_test)

CrossTable(x = particles_test_labels[,1, drop = TRUE], y = particles_test_pred, prop.chisq=FALSE)



```
```{r}
summary(m)
```

# with costs if we specifically care about particle -11


## Train classifier on training data and classify test data

```{r}
#generating classification for trials to find best accuracy

cost_mat <- matrix(c(0, 5000, 5000, 5000, 5000,0,0,0,5000,0,0,0,0,0,0,0), nrow = 4)
#cost_mat

#accuracy = 1:30


#for (i in 1:30){

  m <- C5.0(particles_train, as.factor(particles_train_labels[,1, drop = TRUE]), trials = 30, costs = cost_mat)


  particles_test_pred <- predict(m, particles_test)
    
  #metric
  accuracy[3] <- Accuracy(particles_test_pred, particles_test_labels[,1, drop = TRUE])

#get best accurracy and corresponding index (with respect to Laplace estimator)
#  if (i > 1){
#    if (accuracy[i] > bestacc){
#      bestacc <- accuracy[i]
#      bestk <- i
#    }
#  } else {
#    bestacc <- accuracy[i]
#    bestk <- i
#  }
#}
```
## Metrics

```{r}
#plot(accuracy[1:30])
print(paste("The best accuracy is", accuracy[3]))
```
```{r}

#m <- C5.0(particles_train, as.factor(particles_train_labels[,1, drop = TRUE]), trials = bestk)


#  particles_test_pred <- predict(m, particles_test)

CrossTable(x = particles_test_labels[,1, drop = TRUE], y = particles_test_pred, prop.chisq=FALSE)



```
```{r}
summary(m)
```






# Rules in OneR
```{r}

#m <- OneR(id ~ ., particles[1:45000])


#particles_test_pred <- predict(m, particles_test)
    
  #metric
#accuracy[1] <- Accuracy(particles_test_pred, particles_test_labels[,1, drop = TRUE])


```
```{r}
#print(paste("The accuracy is", accuracy[1]))
```
```{r}



#CrossTable(x = particles_test_labels[,1, drop = TRUE], y = particles_test_pred, prop.chisq=FALSE)



```
```{r}
#summary(m)
```


# Rules in Jrip
```{r}

#m <- Jrip(id ~ ., particles[1:45000])


#particles_test_pred <- predict(m, particles_test)
    
  #metric
#accuracy[2] <- Accuracy(particles_test_pred, particles_test_labels[,1, drop = TRUE])


```
```{r}
#print(paste("The accuracy is", accuracy[2]))
```
```{r}



#CrossTable(x = particles_test_labels[,1, drop = TRUE], y = particles_test_pred, prop.chisq=FALSE)



```
```{r}
#summary(m)
```
